---
title: "reproduce_gsmwlrt_paper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reproduce_gsmwlrt_paper}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(tidyverse)
#library(gsdelayed)
devtools::load_all()
```
## Alternative hypotheses

```{r}

####################################
get_t_pieces = function(t, ts){
  
  if (length(t) > 1) stop("t must be length 1")
  
  pmax(0, pmin(t - ts[-length(ts)], diff(ts)))
}
surv_pieces_simple = function(t, change_points, lambdas){
  
  if (length(t) > 1) stop("t must be length 1")
  if (length(change_points) != length(lambdas) - 1) stop("require one event rate per time period")
  ts = c(0, change_points, Inf)
  
  exp(-sum(get_t_pieces(t, ts) * lambdas))
  
}
####################################
t_seq <- seq(0,24, length.out = 100)

s_c <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                      change_points = c(6), lambdas = log(2) / c(8, 8))
s_e_0 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(0), lambdas = log(2) / c(12.3, 12.3))
s_e_4 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(4), lambdas = log(2) / c(8, 16.6))
s_e_6 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(6), lambdas = log(2) / c(8, 22))


plot(t_seq, s_c, ylim = c(0,1), type = "l", xlab = "Time (months)", ylab = "Survival",
     main = "Alternative hypotheses")
points(t_seq, s_e_0, lty = 1, type = "l", col = 2)
points(t_seq, s_e_4, lty = 1, type = "l", col = 4)
legend("topright", c("control",
                     "delay = 0m",
                     "delay = 4m"), lty = c(1,1,1), col = c(1,2,4))

```



## Reproduce power calculation



```{r}
model_ph <- list(change_points = c(4),
              lambdas_0 = c(log(2) / 8, log(2) / 8),
              lambdas_1 = c(log(2) / 12.3, log(2) / 12.3))

model_nph <- list(change_points = c(4),
                  lambdas_0 = c(log(2) / 8, log(2) / 8),
                  lambdas_1 = c(log(2) / 8, log(2) / 16.6))


recruitment_options = list(list(n_0 = 150, n_1 = 150, r_period = 8, k = 1),
                           list(n_0 = 155, n_1 = 155, r_period = 8, k = 1),
                           list(n_0 = 160, n_1 = 160, r_period = 8, k = 1),
                           list(n_0 = 165, n_1 = 165, r_period = 8, k = 1),
                           list(n_0 = 170, n_1 = 170, r_period = 8, k = 1),
                           list(n_0 = 175, n_1 = 175, r_period = 8, k = 1),
                           list(n_0 = 180, n_1 = 180, r_period = 8, k = 1))


## ph ##
single_stage_options_lrt_ph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 0,
             model = model_ph,
             dco_final = 21,
             alpha_one_sided = 0.025)

pow_ph_0 <- purrr::map_dbl(single_stage_options_lrt_ph, function(x) round(x$overall_power, 2))
pow_ph_0

## nph ##


single_stage_options_lrt_nph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 0,
             model = model_nph,
             dco_final = 21,
             alpha_one_sided = 0.025)

pow_nph_0 <- purrr::map_dbl(single_stage_options_lrt_nph, function(x) round(x$overall_power, 2))
pow_nph_0


## ph ##
single_stage_options_mwlrt6_ph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 6,
             model = model_ph,
             dco_final = 21,
             alpha_one_sided = 0.025)

pow_ph_6 <- purrr::map_dbl(single_stage_options_mwlrt6_ph, function(x) round(x$overall_power, 2))
pow_ph_6

## nph ##

single_stage_options_mwlrt6_nph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 6,
             model = model_nph,
             dco_final = 21,
             alpha_one_sided = 0.025)

pow_nph_6 <- purrr::map_dbl(single_stage_options_mwlrt6_nph, function(x) round(x$overall_power, 2))
pow_nph_6


```

## Reproduce expected duration

```{r}
######################################

recruitment <- list(n_0 = 150, 
                    n_1 = 150, 
                    r_period = 8, 
                    k = 1)

model <- list(change_points = c(4),
              lambdas_0 = c(log(2) / 8, log(2) / 8),
              lambdas_1 = c(log(2) / 8, log(2) / 16.6))

######################################
hsd_obf <- function(t, alpha_one_sided = 0.025, gamma = -4){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
hsd_mid <- function(t, alpha_one_sided = 0.025, gamma = -1.5){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
hsd_poc <- function(t, alpha_one_sided = 0.025, gamma = 1){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}

### - ###
single_stage_design(t_star = 6,
                    model = model,
                    recruitment = recruitment,
                    dco_final = 21,
                    alpha_one_sided = 0.025)[c("overall_power", "n_events")]

### 11 #####
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]

two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]


two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]


### 16 #####
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]

two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]


two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]


### 11 & 16 #####
three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]

three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]


three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]

```



## Reproduce design plots


```{r}
hsd_obf <- function(t, alpha_one_sided = 0.025, gamma = -4){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}

d <- three_stage_design(t_star = 6,
                        model = model,
                        recruitment = recruitment,
                        dco_int_1 = 11,
                        dco_int_2 = 16,
                        dco_final = 21,
                        alpha_spend_f = hsd_obf,
                        alpha_one_sided = 0.025)

######################################
plot(c(11, 16, 21),d$critical_values,
     ylim = c(-4,1),
     xlim = c(0,24),
     xlab = "Time since start of study (months)",
     ylab = "Z statistic",
     main = "Stopping boundaries",
     pch = 16)
abline(h = 0, lty = 2)
lines(x = c(11,11), y = c(-5, d$critical_values[1]))
lines(x = c(16,16), y = c(-5, d$critical_values[2]))
lines(x = c(21,21), y = c(-5, d$critical_values[3]))
######################################
## draw expected number of events

cal_t <- 1:31
exp_n <- numeric(31)
for (i in 1:31){
  exp_n[i] <- single_stage_design(t_star = 6, 
                                  model = model,recruitment = recruitment,
                                  dco = i,
                                  alpha_one_sided = 0.025)$n_events
}

plot(cal_t, exp_n, type = "l", 
     ylim = c(0,300),
     xlim = c(0,32),
     ylab = "Expected n", 
     xlab = "Time since start of study (months)",
     main = "Convert from time to events")

points(c(0,8,31), c(0,300,300), type = "l", lty = 2)
lines(x = c(11,11), y = c(0, d$n_events[1]), lty = 3)
lines(x = c(16,16), y = c(0, d$n_events[2]), lty = 3)
lines(x = c(21,21), y = c(0, d$n_events[3]), lty = 3)

lines(y = c(d$n_events[1],d$n_events[1]), x = c(0, 11), lty = 3)
lines(y = c(d$n_events[2],d$n_events[2]), x = c(0, 16), lty = 3)
lines(y = c(d$n_events[3],d$n_events[3]), x = c(0, 21), lty = 3)


legend(22,150,
       c("Recruited ", "Events"), 
       lty = c(2,1),
       bty = "n")

d$n_events

######################################
plot(d$n_events,d$critical_values,
     ylim = c(-4,1),
     xlim = c(0,220),
     xlab = "Observed number of events",
     ylab = "Z statistic",
     main = "Planned stopping boundaries",
     pch = 16)
abline(h = 0, lty = 2)
lines(x = c(d$n_events[1],d$n_events[1]), y = c(-5, d$critical_values[1]))
lines(x = c(d$n_events[2],d$n_events[2]), y = c(-5, d$critical_values[2]))
lines(x = c(d$n_events[3],d$n_events[3]), y = c(-5, d$critical_values[3]))
lines(x = c(d$n_events[3],d$n_events[3]), y = c(d$critical_values[3], 5), lty = 3)
######################################

```

## Reproduce implementation

```{r}
## simulate a trial

set.seed(45)
dat_uncensored <- gsdelayed::sim_t_uncensored(model = model, 
                                              recruitment = recruitment)
head(dat_uncensored)

#################################################################
dat_interim_1 <- gsdelayed::apply_dco(dat_uncensored, dco = NULL, events = 122)
head(dat_interim_1)

wlrt_interim_1 <-gsdelayed::wlrt(dat_interim_1,
                               trt_colname = "group",
                               time_colname = "time",
                               event_colname = "event",
                               wlr = "mw",
                               t_star = 6)


wlrt_interim_1$v_u / d$var_u[3]
qnorm(hsd_obf(wlrt_interim_1$v_u / d$var_u[3]))


c_1 <- gsdelayed::crit_1_of_3(var_u_int_1 = wlrt_interim_1$v_u,
                       design = d,
                       alpha_spend_f = hsd_obf ,
                       alpha_one_sided = 0.025)

c_1

wlrt_interim_1$z
##############################################################

dat_interim_2 <- gsdelayed::apply_dco(dat_uncensored, dco = NULL, events = 170)
head(dat_interim_2)

wlrt_interim_2 <-gsdelayed::wlrt(dat_interim_2,
                                 trt_colname = "group",
                                 time_colname = "time",
                                 event_colname = "event",
                                 wlr = "mw",
                                 t_star = 6)

wlrt_interim_2$v_u /d$var_u[3]
hsd_obf(wlrt_interim_2$v_u /d$var_u[3])


gsdelayed::crit_2_of_3(var_u_int_2 = wlrt_interim_2$v_u,
                       var_u_int_1 = wlrt_interim_1$v_u,
                       design = d,
                       alpha_spend_f = hsd_obf ,
                       alpha_one_sided = 0.025)

wlrt_interim_2$z


###################################################


dat_interim_3 <- gsdelayed::apply_dco(dat_uncensored, dco = 21)
head(dat_interim_3)

wlrt_interim_3 <-gsdelayed::wlrt(dat_interim_3,
                                 trt_colname = "group",
                                 time_colname = "time",
                                 event_colname = "event",
                                 wlr = "mw",
                                 t_star = 6)
wlrt_interim_3$v_u
d$var_u[3]


gsdelayed::crit_3_of_3(var_u_final = wlrt_interim_3$v_u, 
                       var_u_int_2 = wlrt_interim_2$v_u,
                       var_u_int_1 = wlrt_interim_1$v_u,
                       design = d,
                       alpha_spend_f = hsd_obf ,
                       alpha_one_sided = 0.025)


################################################################
## p-value
gsdelayed::stagewise_p_2(c_1,
                         wlrt_interim_2$z,
                         wlrt_interim_1$v_u, 
                         wlrt_interim_2$v_u )



```



## Double checking

### Comparison with gsDesign

It's possible to compare the boundaries to `gsDesign`

```{r}
### compare 
d$critical_values

### with
gsDesign::gsDesign(k = 3, test.type = 1,
                   timing = d$var_u / d$var_u[3], sfupar = -4)
```


### Comparison with simulation

It's possible to check the power of the (modestly-weighted) log-rank test via simulation.


```{r}

d <- three_stage_design(t_star = 6,
                        model = model,
                        recruitment = recruitment,
                        dco_int_1 = 11,
                        dco_int_2 = 16,
                        dco_final = 21)

system.time({sim_results <- three_stage_sim(n_sims = 100, design = d)})

## compare 
d$overall_power

## with
mean(sim_results$z_1 < sim_results$c_1 | 
       sim_results$z_2 < sim_results$c_2 | 
       sim_results$z_3 < sim_results$c_3)


## compare
d$expected_t

## with 
mean(sim_results$t_1 * (sim_results$z_1 <= sim_results$c_1) + 
     sim_results$t_2 * ((sim_results$z_1 > sim_results$c_1) & 
                          (sim_results$z_2<=sim_results$c_2)) +
     sim_results$t_3 *  ((sim_results$z_1 > sim_results$c_1) & 
                           (sim_results$z_2 > sim_results$c_2)))
```


Let's try another one:



```{r}
d <- three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)


### compare 
d$critical_values

### with
gsDesign::gsDesign(k = 3, test.type = 1, 
                   timing = d$var_u / d$var_u[3], sfupar = -1.5)


sim_results <- three_stage_sim(n_sims = 1000, design = d)

## compare 
d$overall_power

## with
mean(sim_results$z_1 < sim_results$c_1 |
       sim_results$z_2 < sim_results$c_2 | 
       sim_results$z_3 < sim_results$c_3)


## compare
d$expected_t

## with 
mean(sim_results$t_1 * (sim_results$z_1 <= sim_results$c_1) + 
     sim_results$t_2 *  ((sim_results$z_1 > sim_results$c_1) &
                           (sim_results$z_2 <= sim_results$c_2)) +
     sim_results$t_3 *  ((sim_results$z_1 > sim_results$c_1) &
                           (sim_results$z_2 > sim_results$c_2)))
```


### Mis-specified design assumptions


```{r}

####################################
get_t_pieces = function(t, ts){
  
  if (length(t) > 1) stop("t must be length 1")
  
  pmax(0, pmin(t - ts[-length(ts)], diff(ts)))
}

surv_pieces_simple = function(t, change_points, lambdas){
  
  if (length(t) > 1) stop("t must be length 1")
  if (length(change_points) != length(lambdas) - 1) stop("require one event rate per time period")
  ts = c(0, change_points, Inf)
  
  exp(-sum(get_t_pieces(t, ts) * lambdas))
  
}
####################################
t_seq <- seq(0,24, length.out = 100)

s_c <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                      change_points = c(6), lambdas = log(2) / c(8, 8))
s_e_0 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(0), lambdas = log(2) / c(12.3, 12.3))
s_e_4 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(4), lambdas = log(2) / c(8, 16.6))
s_e_8 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(8), lambdas = log(2) / c(8, 35))


plot(t_seq, s_c, ylim = c(0,1), type = "l", xlab = "Time (months)", ylab = "Survival",
     main = "Alternative hypotheses")
points(t_seq, s_e_0, lty = 1, type = "l", col = 2)
points(t_seq, s_e_4, lty = 1, type = "l", col = 4)
points(t_seq, s_e_8, lty = 1, type = "l", col = 6)
legend("topright", c("control",
                     "delay = 0m",
                     "delay = 4m"), lty = c(1,1,1), col = c(1,2,4))

```

```{r}
t_seq <- seq(0,24, length.out = 100)

s_c <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                      change_points = c(6), lambdas = log(2) / c(6, 6))
s_e_0 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(0), lambdas = log(2) / c(9.2, 9.2))
s_e_4 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(4), lambdas = log(2) / c(6, 12.6))
s_e_8 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(8), lambdas = log(2) / c(6, 25))


plot(t_seq, s_c, ylim = c(0,1), type = "l", xlab = "Time (months)", ylab = "Survival",
     main = "Alternative hypotheses")
points(t_seq, s_e_0, lty = 1, type = "l", col = 2)
points(t_seq, s_e_4, lty = 1, type = "l", col = 4)
points(t_seq, s_e_8, lty = 1, type = "l", col = 6)
legend("topright", c("control",
                     "delay = 0m",
                     "delay = 4m"), lty = c(1,1,1), col = c(1,2,4))
```


```{r}
t_seq <- seq(0,24, length.out = 100)

s_c <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                      change_points = c(6), lambdas = log(2) / c(10, 10))
s_e_0 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(0), lambdas = log(2) / c(15.4, 15.4))
s_e_4 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(4), lambdas = log(2) / c(10, 21.6))
s_e_8 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(8), lambdas = log(2) / c(10, 45))


plot(t_seq, s_c, ylim = c(0,1), type = "l", xlab = "Time (months)", ylab = "Survival",
     main = "Alternative hypotheses")
points(t_seq, s_e_0, lty = 1, type = "l", col = 2)
points(t_seq, s_e_4, lty = 1, type = "l", col = 4)
points(t_seq, s_e_8, lty = 1, type = "l", col = 6)
legend("topright", c("control",
                     "delay = 0m",
                     "delay = 4m"), lty = c(1,1,1), col = c(1,2,4))
```


## Corresponding standard log rank tests

```{r}


################################################################
model_8m_control_null <- list(change_points = c(4),
                              lambdas_0 = c(log(2) / 8, log(2) / 8),
                              lambdas_1 = c(log(2) / 8, log(2) / 8))


model_8m_control_0m_delay <- list(change_points = c(4),
                                  lambdas_0 = c(log(2) / 8, log(2) / 8),
                                  lambdas_1 = c(log(2) / 12.3, log(2) / 12.3))


model_8m_control_4m_delay <- list(change_points = c(4),
                                  lambdas_0 = c(log(2) / 8, log(2) / 8),
                                  lambdas_1 = c(log(2) / 8, log(2) / 16.6))



model_8m_control_8m_delay <- list(change_points = c(8),
                                  lambdas_0 = log(2) / c(8,8),
                                  lambdas_1 = log(2) / c(8, 35))


############################################################
model_6m_control_null <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(6,6),
                                  lambdas_1 = log(2) / c(6, 6))

model_6m_control_0m_delay <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(6,6),
                                  lambdas_1 = log(2) / c(9.2,9.2))

model_6m_control_4m_delay <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(6,6),
                                  lambdas_1 = log(2) / c(6, 12.6))

model_6m_control_8m_delay <- list(change_points = c(8),
                                  lambdas_0 = log(2) / c(6,6),
                                  lambdas_1 = log(2) / c(6, 25))


############################################################
model_10m_control_null <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(10,10),
                                  lambdas_1 = log(2) / c(10, 10))

model_10m_control_0m_delay <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(10,10),
                                  lambdas_1 = log(2) / c(15.4,15.4))

model_10m_control_4m_delay <- list(change_points = c(4),
                                  lambdas_0 = log(2) / c(10,10),
                                  lambdas_1 = log(2) / c(10,21.6))


model_10m_control_8m_delay <- list(change_points = c(8),
                                  lambdas_0 = log(2) / c(10,10),
                                  lambdas_1 = log(2) / c(10, 45))


```



## Mis-specified recruitment

```{r}
recruitment_1 <- list(n_0 = 150, 
                      n_1 = 150, 
                      r_period = 8, 
                      k = 1)


recruitment_2 <- list(n_0 = 150, 
                      n_1 = 150, 
                      r_period = 8, 
                      k = 2)



recruitment_3 <- list(n_0 = 150, 
                      n_1 = 150, 
                      r_period = 15, 
                      k = 2)


```



## Designs


```{r}
d_8m_control_4m_delay <- three_stage_design(t_star = 6,
                                            model = model_8m_control_4m_delay,
                                            recruitment = recruitment_1,
                                            dco_int_1 = 11,
                                            dco_int_2 = 16,
                                            dco_final = 21,
                                            alpha_spend_f = hsd_mid,
                                            alpha_one_sided = 0.025)

d_8m_control_4m_delay_lr <- three_stage_design(t_star = 0,
                                               model = model_8m_control_4m_delay,
                                               recruitment = recruitment_1,
                                               dco_int_1 = 11,
                                               dco_int_2 = 16,
                                               dco_final = 21,
                                               alpha_spend_f = hsd_mid,
                                               alpha_one_sided = 0.025)


#################################################
d_6m_control_4m_delay <- three_stage_design(t_star = 6,
                                            model = model_6m_control_4m_delay,
                                            recruitment = recruitment_1,
                                            dco_int_1 = 11,
                                            dco_int_2 = 16,
                                            dco_final = 21,
                                            alpha_spend_f = hsd_mid,
                                            alpha_one_sided = 0.025)


```



## Simulations

```{r}
sim_results <- three_stage_sim(n_sims = 10, design = d_8m_control_4m_delay)
```























