---
title: "Reproduce output from paper"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{reproduce_vignette}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE,
                      comment =  "#>")
```

```{r setup}
#library(gsdelayed)
devtools::load_all(".")

library(tidyverse)
library(scales)
library(survival)
library(survminer)
library(survRM2)
library(gsDesign)
```

## Alternative hypotheses

```{r}
####################################
get_t_pieces = function(t, ts){
  
  if (length(t) > 1) stop("t must be length 1")
  
  pmax(0, pmin(t - ts[-length(ts)], diff(ts)))
}
surv_pieces_simple = function(t, change_points, lambdas){
  
  if (length(t) > 1) stop("t must be length 1")
  if (length(change_points) != length(lambdas) - 1) stop("require one event rate per time period")
  ts = c(0, change_points, Inf)
  
  exp(-sum(get_t_pieces(t, ts) * lambdas))
  
}
####################################
t_seq <- seq(0,24, length.out = 100)
s_c <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                      change_points = c(6), lambdas = log(2) / c(8, 8))
s_e_0 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(0), lambdas = log(2) / c(12.3, 12.3))
s_e_4 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(4), lambdas = log(2) / c(8, 16.6))
s_e_6 <- purrr::map_dbl(t_seq, surv_pieces_simple, 
                        change_points = c(6), lambdas = log(2) / c(8, 22))

df_alternative = data.frame(x = rep(t_seq,3),
                            y = c(s_c,s_e_0,s_e_4),
                            type = c(rep("Control",length(s_c)),
                                     rep("Delay = 0 months",length(s_c)),
                                     rep("Delay = 4 months",length(s_c))))

ggplot(df_alternative, aes(x = x, y = y, color = type)) + 
  geom_line(size = 1) +
  scale_y_continuous("Survival",limits = c(0,1)) +
  scale_x_continuous("Time (months)") +
  labs(color = "", title = "Alternative hypotheses") +
  theme_bw() +
  theme(text = element_text(size = 16),
        legend.position = c(0.75,0.9),
        #legend.text = element_text(face = "italic"),
        legend.background = element_rect(fill="transparent"),
        legend.key = element_rect(fill="transparent"))
```

## Reproduce power calculation



```{r}
model_ph <- list(change_points = c(4),
              lambdas_0 = c(log(2) / 8, log(2) / 8),
              lambdas_1 = c(log(2) / 12.3, log(2) / 12.3))
model_nph <- list(change_points = c(4),
                  lambdas_0 = c(log(2) / 8, log(2) / 8),
                  lambdas_1 = c(log(2) / 8, log(2) / 16.6))
recruitment_options = list(list(n_0 = 150, n_1 = 150, r_period = 8, k = 1),
                           list(n_0 = 155, n_1 = 155, r_period = 8, k = 1),
                           list(n_0 = 160, n_1 = 160, r_period = 8, k = 1),
                           list(n_0 = 165, n_1 = 165, r_period = 8, k = 1),
                           list(n_0 = 170, n_1 = 170, r_period = 8, k = 1),
                           list(n_0 = 175, n_1 = 175, r_period = 8, k = 1),
                           list(n_0 = 180, n_1 = 180, r_period = 8, k = 1))
## ph ##
single_stage_options_lrt_ph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 0,
             model = model_ph,
             dco_final = 21,
             alpha_one_sided = 0.025)
pow_ph_0 <- purrr::map_dbl(single_stage_options_lrt_ph, function(x) round(x$overall_power, 2))
pow_ph_0
## nph ##
single_stage_options_lrt_nph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 0,
             model = model_nph,
             dco_final = 21,
             alpha_one_sided = 0.025)
pow_nph_0 <- purrr::map_dbl(single_stage_options_lrt_nph, function(x) round(x$overall_power, 2))
pow_nph_0
## ph ##
single_stage_options_mwlrt6_ph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 6,
             model = model_ph,
             dco_final = 21,
             alpha_one_sided = 0.025)
pow_ph_6 <- purrr::map_dbl(single_stage_options_mwlrt6_ph, function(x) round(x$overall_power, 2))
pow_ph_6
## nph ##
single_stage_options_mwlrt6_nph <- recruitment_options %>% 
  purrr::map(gsdelayed::single_stage_design,
             t_star = 6,
             model = model_nph,
             dco_final = 21,
             alpha_one_sided = 0.025)
pow_nph_6 <- purrr::map_dbl(single_stage_options_mwlrt6_nph, function(x) round(x$overall_power, 2))
pow_nph_6
```

## Reproduce expected duration

```{r}
######################################
recruitment <- list(n_0 = 150, 
                    n_1 = 150, 
                    r_period = 8, 
                    k = 1)
model <- list(change_points = c(4),
              lambdas_0 = c(log(2) / 8, log(2) / 8),
              lambdas_1 = c(log(2) / 8, log(2) / 16.6))
######################################
hsd_obf <- function(t, alpha_one_sided = 0.025, gamma = -4){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
hsd_mid <- function(t, alpha_one_sided = 0.025, gamma = -1.5){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
hsd_poc <- function(t, alpha_one_sided = 0.025, gamma = 1){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
### - ###
single_stage_design(t_star = 6,
                    model = model,
                    recruitment = recruitment,
                    dco_final = 21,
                    alpha_one_sided = 0.025)[c("overall_power", "n_events")]
### 11 #####
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 11,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
### 16 #####
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
two_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
### 11 & 16 #####
three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_obf,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_mid,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
three_stage_design(t_star = 6,
                 model = model,
                 recruitment = recruitment,
                 dco_int_1 = 11,
                 dco_int_2 = 16,
                 dco_final = 21,
                 alpha_spend_f = hsd_poc,
                 alpha_one_sided = 0.025)[c("overall_power", "expected_t")]
```


## Reproduce design plots


```{r}
hsd_obf <- function(t, alpha_one_sided = 0.025, gamma = -4){
  alpha_one_sided * (1 - exp(-gamma * t)) / (1 - exp(-gamma))}
d <- three_stage_design(t_star = 6,
                        model = model,
                        recruitment = recruitment,
                        dco_int_1 = 11,
                        dco_int_2 = 16,
                        dco_final = 21,
                        alpha_spend_f = hsd_obf,
                        alpha_one_sided = 0.025)

d$n_events
d$critical_values


df_critical_values = data.frame (x = d$n_events,
                                 y = d$critical_values)

ggplot(df_critical_values,aes(x = x, y = y)) +
  geom_point(size = 3) +
  scale_x_continuous("Observed number of events", limits = c(0,220)) +
  scale_y_continuous("Z statistic", limits = c(-4,1)) +
  labs(title = "Planned stopping boundaries") +
  geom_segment(aes(x = d$n_events[1], y = -4, xend = d$n_events[1], yend = d$critical_values[1]), color = "black", size = 0.75) +
  geom_segment(aes(x = d$n_events[2], y = -4, xend = d$n_events[2], yend = d$critical_values[2]), color = "black", size = 0.75) +
  geom_segment(aes(x = d$n_events[3], y = -4, xend = d$n_events[3], yend = d$critical_values[3]), color = "black", size = 0.75) +
  geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.75) +
  theme_bw() +
  theme(text = element_text(size = 16),
        legend.position = c(0.85,0.2),
        legend.background = element_rect(fill="transparent"),
        legend.key = element_rect(fill="transparent"))

######################################
## draw expected number of events
cal_t <- 1:31
exp_n <- numeric(31)
for (i in 1:31){
  exp_n[i] <- single_stage_design(t_star = 6, 
                                  model = model,recruitment = recruitment,
                                  dco = i,
                                  alpha_one_sided = 0.025)$n_events
}

df_event_times = data.frame(x = c(0,8,31,cal_t), 
                            y = c(0,300,300,exp_n),
                            type = c("Recruited","Recruited","Recruited",rep("Events",length(cal_t))))

df_event_times$type = factor(df_event_times$type,
                             levels = c("Recruited","Events"), 
                             labels = c("Recruited","Events"),
                             ordered = TRUE)

df_dotted_lines1 = data.frame(x = c(0,11,11),y = c(d$n_events[1],d$n_events[1],0))
df_dotted_lines2 = data.frame(x = c(0,16,16),y = c(d$n_events[2],d$n_events[2],0))
df_dotted_lines3 = data.frame(x = c(0,21,21),y = c(d$n_events[3],d$n_events[3],0))

colors_ggplot_2 = hue_pal()(2)

ggplot(df_event_times,aes(x = x, y = y, color = type)) +
  geom_line(df_dotted_lines1, mapping = aes(x = x, y = y), colour = "black", linetype = "dashed") +
  geom_line(df_dotted_lines2, mapping = aes(x = x, y = y), colour = "black", linetype = "dashed") +
  geom_line(df_dotted_lines3, mapping = aes(x = x, y = y), colour = "black", linetype = "dashed") +
  geom_line(size = 1) +
  scale_color_manual(values=colors_ggplot_2) +
  scale_x_continuous("Time since start of study (months)", limits = c(0,32)) +
  scale_y_continuous("Expected n", limits = c(0,300)) +
  labs(color = " ", title = "Convert from time to events") +
  theme_bw() +
  theme(text = element_text(size = 16),
        legend.position = c(0.85,0.2),
        legend.background = element_rect(fill="transparent"),
        legend.key = element_rect(fill="transparent"))

```




## Reproduce implementation

```{r}
## simulate a trial
set.seed(45)
dat_uncensored <- gsdelayed::sim_t_uncensored(model = list(change_points = c(4),
                                                           lambdas_0 = c(0.087, 0.087),
                                                           lambdas_1 = c(0.087, 0.055)), 
                                              recruitment = recruitment)
head(dat_uncensored)
#################################################################
dat_interim_1 <- gsdelayed::apply_dco(dat_uncensored, dco = NULL, events = 122)
head(dat_interim_1)


km_IA1 <- survival::survfit(survival::Surv(time, event) ~ group,
                     data = dat_interim_1)

p_km_1 <- survminer::ggsurvplot(km_IA1,
                      data = dat_interim_1,
                      legend.labs = c("Control","Experimental"),
                      risk.table = TRUE,
                      conf.int = TRUE,
                      surv.median.line = "hv",
                      title = "A) Interim Analysis 1",
                      break.x.by = 2,
                      legend.title = "",
                      xlab = "Time (months)",
                      ylab = "Overall survival",
                      risk.table.fontsize = 4)

wlrt_interim_1 <-gsdelayed::wlrt(dat_interim_1,
                               trt_colname = "group",
                               time_colname = "time",
                               event_colname = "event",
                               wlr = "mw",
                               t_star = 6)

wlrt_interim_1$v_u / d$var_u[3]

qnorm(hsd_obf(wlrt_interim_1$v_u / d$var_u[3]))

c_1 <- gsdelayed::crit_1_of_3(info_frac_sf = wlrt_interim_1$v_u / d$var_u[3],
                              alpha_spend_f = hsd_obf,
                              alpha_one_sided = 0.025)
c_1

wlrt_interim_1$u
wlrt_interim_1$v_u
wlrt_interim_1$z


c_1_star <- gsdelayed::crit_1_of_3(info_frac_sf = wlrt_interim_1$v_u / d$var_u[3],
                                   alpha_spend_f = function(t, alpha_one_sided) (d$alpha_spend_f(d$var_u / max(d$var_u)))[1],
                                   alpha_one_sided = 0.025)
c_1_star
```


```{r}
### interim 2
##############################################################
dat_interim_2 <- gsdelayed::apply_dco(dat_uncensored, dco = NULL, events = 170)
head(dat_interim_2)


km_IA2 <- survival::survfit(survival::Surv(time, event) ~ group,
                            data = dat_interim_2)

p_km_2 <- survminer::ggsurvplot(km_IA2,
                      data = dat_interim_2,
                      legend.labs = c("Control","Experimental"),
                      risk.table = TRUE,
                      conf.int = TRUE,
                      surv.median.line = "hv",
                      title = "B) Interim Analysis 2",
                      break.x.by = 2,
                      legend.title = "",
                      xlab = "Time (months)",
                      ylab = "",
                      risk.table.fontsize = 4)


wlrt_interim_2 <-gsdelayed::wlrt(dat_interim_2,
                                 trt_colname = "group",
                                 time_colname = "time",
                                 event_colname = "event",
                                 wlr = "mw",
                                 t_star = 6)

wlrt_interim_2$v_u /d$var_u[3]

hsd_obf(wlrt_interim_2$v_u /d$var_u[3])

c_2 <- gsdelayed::crit_2_of_3(info_frac_sf = wlrt_interim_2$v_u / d$var_u[3],
                              info_frac_1_2 =  wlrt_interim_1$v_u / wlrt_interim_2$v_u,
                              crit_1 = c_1,
                              alpha_spend_f = hsd_obf ,
                              alpha_one_sided = 0.025)

c_2 



c_2_star <- gsdelayed::crit_2_of_3(info_frac_sf = wlrt_interim_2$v_u / d$var_u[3],
                                   info_frac_1_2 =  wlrt_interim_1$v_u / wlrt_interim_2$v_u,
                                   crit_1 = c_1_star,
                                   alpha_spend_f = function(t, alpha_one_sided) (d$alpha_spend_f(d$var_u / max(d$var_u)))[2],
                                   alpha_one_sided = 0.025)

c_2_star 

hsd_obf(wlrt_interim_2$v_u /d$var_u[3])

wlrt_interim_2$z
wlrt_interim_2$u
wlrt_interim_2$v_u

```



```{r}
### final analysis
###################################################
dat_interim_3 <- gsdelayed::apply_dco(dat_uncensored, dco = NULL, events = 203)
head(dat_interim_3)




km_IA3 <- survival::survfit(survival::Surv(time, event) ~ group,
                            data = dat_interim_3)

p_km_3 <- survminer::ggsurvplot(km_IA3,
                      data = dat_interim_3,
                      legend.labs = c("Control","Experimental"),
                      risk.table = TRUE,
                      conf.int = TRUE,
                      surv.median.line = "hv",
                      title = "C) Final Analysis",
                      break.x.by = 2,
                      legend.title = "",
                      xlab = "Time (months)",
                      ylab = "",
                      risk.table.fontsize = 4)






wlrt_interim_3 <-gsdelayed::wlrt(dat_interim_3,
                                 trt_colname = "group",
                                 time_colname = "time",
                                 event_colname = "event",
                                 wlr = "mw",
                                 t_star = 6)
wlrt_interim_3$u
wlrt_interim_3$v_u
wlrt_interim_3$z

d$var_u[3]

c_3 <- gsdelayed::crit_3_of_3(info_frac_1_3 = wlrt_interim_1$v_u / wlrt_interim_3$v_u,
                       info_frac_2_3 = wlrt_interim_2$v_u / wlrt_interim_3$v_u,
                       crit_1 = c_1,
                       crit_2 = c_2,
                       alpha_spend_f = hsd_obf,
                       alpha_one_sided = 0.025)

c_3


c_3_star <- gsdelayed::crit_3_of_3(info_frac_1_3 = wlrt_interim_1$v_u / wlrt_interim_3$v_u,
                       info_frac_2_3 = wlrt_interim_2$v_u / wlrt_interim_3$v_u,
                       crit_1 = c_1_star,
                       crit_2 = c_2_star,
                       alpha_spend_f = function(t, alpha_one_sided) (d$alpha_spend_f(d$var_u / max(d$var_u)))[],
                       alpha_one_sided = 0.025)


c_3_star

```



```{r}
## plotting and inference
lrt_1 <-gsdelayed::wlrt(dat_interim_1,
                        trt_colname = "group",
                        time_colname = "time",
                        event_colname = "event",
                        wlr = "mw",
                        t_star = 0)


lrt_2 <-gsdelayed::wlrt(dat_interim_2,
                        trt_colname = "group",
                        time_colname = "time",
                        event_colname = "event",
                        wlr = "mw",
                        t_star = 0)


lrt_3 <-gsdelayed::wlrt(dat_interim_3,
                        trt_colname = "group",
                        time_colname = "time",
                        event_colname = "event",
                        wlr = "mw",
                        t_star = 0)


df_IA3 = data.frame(x = c(d$n_events[1],d$n_events[2],d$n_events[3]),
                    y = c(wlrt_interim_1$z, wlrt_interim_2$z, wlrt_interim_3$z),
                    y2 = c(lrt_1$z, lrt_2$z, lrt_3$z))

p_bounds <- ggplot(df_critical_values,aes(x = x, y = y)) +
  geom_point(size = 3) +
  scale_x_continuous("Observed number of events", limits = c(0,220), 
                     breaks = c(0,122,170, 203)) +
  scale_y_continuous("Z statistic", limits = c(-4,1)) +
  labs(title = "Planned stopping boundaries") +
  geom_segment(aes(x = d$n_events[1], y = -4, xend = d$n_events[1], yend = d$critical_values[1]), color = "black", size = 0.75) +
  geom_segment(aes(x = d$n_events[2], y = -4, xend = d$n_events[2], yend = d$critical_values[2]), color = "black", size = 0.75) +
  geom_segment(aes(x = d$n_events[3], y = -4, xend = d$n_events[3], yend = d$critical_values[3]), color = "black", size = 0.75) +
  geom_hline(yintercept=0, linetype="dashed", color = "black", size=0.75) +
  theme_bw() +
  geom_point(data = df_IA3, mapping = aes(x = x, y = y), shape = 4, size = 3, colour = "blue") +
  geom_point(data = df_IA3, mapping = aes(x = x, y = y2), shape = 2, size = 3, colour = "red") +
  geom_line(df_IA3, mapping = aes(x = x, y = y), size = 0.25, linetype = "dashed", colour = "blue") +
  theme(text = element_text(size = 16),
        legend.position = c(0.85,0.2),
        #legend.text = element_text(face = "italic"),
        legend.background = element_rect(fill="transparent"),
        legend.key = element_rect(fill="transparent"))


survminer::arrange_ggsurvplots(list(p_km_1, p_km_3, p_km_2), ncol = 2, nrow = 2)

################################################################
## p-value
gsdelayed::stagewise_p_3(c_1_star,
                         c_2_star,
                         wlrt_interim_3$z,
                         wlrt_interim_1$v_u, 
                         wlrt_interim_2$v_u,
                         wlrt_interim_3$v_u)


#Survival probabilities

surv_table = summary(km_IA3)

surv_table_control <- data.frame(group = "Control",
                         time = surv_table$time[surv_table$strata == "group=control"],
                         n.risk = surv_table$n.risk[surv_table$strata == "group=control"],
                         n.event = surv_table$n.event[surv_table$strata == "group=control"],
                         n.censor = surv_table$n.censor[surv_table$strata == "group=control"],
                         surv = surv_table$surv[surv_table$strata == "group=control"],
                         upper = surv_table$upper[surv_table$strata == "group=control"],
                         lower = surv_table$lower[surv_table$strata == "group=control"])

head(surv_table_control)


surv_table_experimental <- data.frame(group = "Experimental",
                                 time = surv_table$time[surv_table$strata == "group=experimental"],
                                 n.risk = surv_table$n.risk[surv_table$strata == "group=experimental"],
                                 n.event = surv_table$n.event[surv_table$strata == "group=experimental"],
                                 n.censor = surv_table$n.censor[surv_table$strata == "group=experimental"],
                                 surv = surv_table$surv[surv_table$strata == "group=experimental"],
                                 upper = surv_table$upper[surv_table$strata == "group=experimental"],
                                 lower = surv_table$lower[surv_table$strata == "group=experimental"])

head(surv_table_experimental)

#Restricted mean survival time
rmst2(dat_interim_3$time, as.numeric(dat_interim_2$event), ifelse(dat_interim_2$group == "control",0,1), tau = 18)

#Quantiles of the survival
round(quantile(km_IA2, c(0.01, 0.05, 0.1, 0.25, 0.5, 0.75))$quantile,3)

```




## Double checking

### Comparison with gsDesign

It's possible to compare the boundaries to `gsDesign`

```{r}
### compare 
d$critical_values
### with
gsDesign::gsDesign(k = 3, test.type = 1,
timing = d$var_u / d$var_u[3], sfupar = -4)
```


### Comparison with simulation

It's possible to check the power of the (modestly-weighted) log-rank test via simulation.


```{r}
d <- three_stage_design(t_star = 6,
                        model = model,
                        recruitment = recruitment,
                        dco_int_1 = 11,
                        dco_int_2 = 16,
                        dco_final = 21)
sim_results <- three_stage_sim(n_sims = 1000, design = d)
## compare 
d$overall_power
## with
mean(sim_results$z_1 < sim_results$c_1 | 
       sim_results$z_2 < sim_results$c_2 | 
       sim_results$z_3 < sim_results$c_3)
## compare
d$expected_t
## with 
mean(sim_results$t_1 * (sim_results$z_1 <= sim_results$c_1) + 
       sim_results$t_2 * ((sim_results$z_1 > sim_results$c_1) & 
                            (sim_results$z_2<=sim_results$c_2)) +
       sim_results$t_3 *  ((sim_results$z_1 > sim_results$c_1) & 
                             (sim_results$z_2 > sim_results$c_2)))
```


Let's try another one:

```{r}
d <- three_stage_design(t_star = 6,
model = model,
recruitment = recruitment,
dco_int_1 = 11,
dco_int_2 = 16,
dco_final = 21,
alpha_spend_f = hsd_mid,
alpha_one_sided = 0.025)
### compare 
d$critical_values
### with
gsDesign::gsDesign(k = 3, test.type = 1, 
timing = d$var_u / d$var_u[3], sfupar = -1.5)
sim_results <- three_stage_sim(n_sims = 1000, design = d)
## compare 
d$overall_power
## with
mean(sim_results$z_1 < sim_results$c_1 |
sim_results$z_2 < sim_results$c_2 | 
sim_results$z_3 < sim_results$c_3)
## compare
d$expected_t
## with 
mean(sim_results$t_1 * (sim_results$z_1 <= sim_results$c_1) + 
sim_results$t_2 *  ((sim_results$z_1 > sim_results$c_1) &
(sim_results$z_2 <= sim_results$c_2)) +
sim_results$t_3 *  ((sim_results$z_1 > sim_results$c_1) &
(sim_results$z_2 > sim_results$c_2)))
```
